The Galactic Chemical Evolution of Carbon: Implications for Stellar Nucleosynthesis

### 2023-07-25 carbon_paper/tex/main.tex
more consise draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib]{mnras}

@@

% citations
\defcitealias{james+21}{J21}
\newcommand{\JJ}{\citetalias{james+21}}
\newcommand{\VICE}{\textsc{vice}}
\newcommand{\citetjack}{Roberts et al.~(2023, in prep.)}
\newcommand{\citepjack}{(Roberts et al.~2023, in prep.)}
\newcommand{\citealtjack}{Robert et al.~2023, in prep.}
\newcommand{\cxi}{\texttt{\hyperlink{C11}{C11}}}
\newcommand{\kx}{\texttt{\hyperlink{K10}{K10}}}
\newcommand{\kxvi}{\texttt{\hyperlink{K16}{K16}}}
\newcommand{\vxiii}{\texttt{\hyperlink{V13}{V13}}}

@@

% other
\makeatletter
\newcommand{\C}[1][\@nil]{
    \def\tmp{#1}%
    \ifx\tmp\@nnil%
        \ensuremath{\rm C}%
    \else%
        \ifmmode ^{#1}{\rm C}%
        \else $^{#1}$C%
        \fi%
\fi }
\makeatother

\newcommand{\Yct}{\ensuremath{y_{\rm C}}}
\newcommand{\Ycc}{\ensuremath{y_{\rm C}^{\rm cc}}}
\newcommand{\Yoc}{\ensuremath{y_{\rm Mg}^{\rm cc}}}
\newcommand{\Ycagb}{\ensuremath{y_{\rm C}^{\rm agb}}}
\newcommand{\y}{\ensuremath{\rotatebox[origin=B,y=0.5ex]{180}{y}}}

\newcommand{\Mo}{%
    \ifmmode {\rm M}_{\sun}%
    \else M$_{\sun}$
    \fi}
\newcommand{\Zo}{%
    \ifmmode Z_{\sun}%
    \else $Z_{\sun}$%
    \fi}

@@

% Abstract of the paper
\begin{abstract}
% context
C is an important element across astronomy; however, its origin remains poorly understood. 
% 
We aim to constrain the stellar yields of C through multi-zone Galactic chemical evolution models by comparing predictions with \apogee\ subgiants abundances.
% 
We find that \caafe\ is an emirical estimate of the delayed C sources, enabling us to estimate that \agb\ stars and \cc\ produce about 20\% and 80\% of C, respectively.  
The \caah\ trend instead represents the equilibrium abundances of C and Mg. 
We use the \caah\ trend to estimate the CCSNe C/Mg yield, determining that  $\Ycc/\Yoc = EQUATION$ when including \agb\ C. 
% misc
Our models are relatively independent of uniform scaling of yields and outflows, and alternate star formation histories. 
However, the stars which contribute to \agb\ C production and the \ia\ delay time distribution of Fe contribute uncertainties to our conclusions. 
% gas phase
While reliable gas-phase and low-metallicity measurments of C are challenging, we find that our model and a singlezone model with our recommended yields replicate the broad trends of \caah across different environments and metallicities.

@@

Carbon is a distinctive and well-studied element in astronomy. 
Formed in the cores of stars during He fusion, C is the lightest directly synthesized element after He, one of the only light elements formed in low-mass stars, and one the most abundant metals \citep[e.g.][]{jennifer19, KL14}.\footnotemark{}
C structurally changes the environments it pollutes---regulating stellar evolution, facilitating the formation of stars and planets, and forming the basis of earthly life.
% C/N as age indicator?
As such, understanding the origin of C has wide ranging implications. 
While we know both lower-intermediate-mass and high-mass stars produce C, the relative importance of each process is still unknown.

@@

Galactic chemical evolution (\gce) is a powerful tool, capable of uncovering the origins of the elements. 
Each enrichment process has characteristic chemical signatures and timescales, enabling us to reconstruct chemical histories.
Many previous works have used GCE in attemt to understand C abundances, whether through using theoretical stellar models \citep{DTS78, prantzos+18, chiappini+03} 
or understanding  observation \citep{tinsley79, HEK00, BF06, rybizki+17, berg+19, KKL20}.
(See also review in \citealt{romano22}.)
%
Every study agrees that C is produced by a combination of high-mass and \lims{}, different studies disagree on which process is dominant. 
For example CITES conclude \hms{} dominate wherease CITES conclude \lims{} contribute the majority of C.
C is also generally understood to have strong metallicity dependent CCSNe enrichment.

One of the primary uncertainties of GCE models are nucleosynthetic yields. Yield predictions -- the amounts of each chemical element stars produce --
are shaped by poorly understood processes, including mass loss, nuclear reaction rates, rotational mixing, convection, and explodability \citep{romano+10,KL14,ventura+13, LC18, emily+21}.
To better understand where C comes from and how it evolves, our aim is to combine \apogee\ observations and multi-zone models to develop observationally-consistant yields.
\cite{james+23} examined similar \gce{} models of N (which is closely related to C), finding that trends in N and O are explained by the metallicity dependence of N/O yields. \citet{james+23} determine that \agb\ N abundances roughly dependence on metallicity (i.e. $y_{\rm N}/y_{\rm O} \propto Z$). 
Here, we extend their models to C, deriving similar constraints on C/Mg yields. We assess which yield prescriptions reproduce Galactic abundance trends while investigating the impact of \gce{} model assumptions, such as the star formation history (\sfh{}) and outflow mass loading.

% \footnotetext{e.g. \citet{HEK00,PVT10,berg+12, berg+20, skillman+20, izotov+12, james2+15, dopita+16}.} 
We know, from observations of MW stars and extragalactic gas measurments, that C/O traces a banana shape in O/H (see Fig.~\ref{fig:gas_phase} and section~\ref{sec:gas}). 
C (and O) have been observed in local and extragalactic HII regions (CITES), dwarf galaxies \citep[e.g.]{berg+19}, halo and thick disc Milky Way stars \cite[e.g.][]{fabbian+09, nissen+14, lambert81, laird85, lambert86}, and damped Lyman-alpha systems. 
\footnote{DLAs; high redshift clouds of gas observed as absorption lines in yet higher redshift quasars}
The many observations of C and O paint a broadly consistant picture of a global trend.
At the very lowest metallicities, where $[{\rm O/H}]\lesssim-2$,\footnotemark{} observations suggest that C/O declines with metallicity.
From $-2 \lesssim {\rm [O/H]}\lesssim -1$, C/O is roughtly constant with metallicities. 
And, at higher metallicities $-1 \lesssim {\rm [O/H]}$, C/O increases with increasing metallicity.

\footnotetext{In this paper, we use the standard notation for chemical abundances. $[A/B] = \log_{10}\left(A/B\right) - \log_{10}\left(A_{\sun}/B_{\sun}\right)$, i.e. $[A/B]$ is the logarithm of the ratio between A and B, scaled such that $[A/B]=0$ for the sun. Solar abundances are as measured in \citet{asplund+09}.}

However, observations of C are not without challenges. 
As stars ascend the red giant branch (RGB) at the end of H core burning, C abundances are modified through \textit{first dredge up}\footnotemark{}. Because the effects of first dredge up are mass-dependent, C and N abundances are frequently used as age indicators for RGB stars \citep{MG15, martig16, hasselquist19, vincenzo+21}. Evolved RGB stars thus require modelling-dependent corrections to recover birth C and N abundances \citep[e.g.][]{vincenzo+21}. 
C is also challenging to determine in optical spectroscopy. 
C lines are relatively faint, and surveys such as GALAH struggle against low detection rates, potentially biasing sample measurments. 
(e.g. gaia-eso AS WELL \citealt{franchini+20}).
In the gas phase, HII regions are our best window into C abundances. Unfortunantly, C lacks strong collisional excitation lines, and recombination lines fall in the ultraviolet without nearby reference H lines \citep{skillman+20}. Recombination-line and collisional-excitation-line measurements furthermore disagree by a factor of \about{2} \citep{GR07}.

\footnotetext{First dredge up is when material from CNO processed core is mixed with the (convective) stellar envelop during the ascent onto the RGB. As a result, C is decreased and N is increased depending on the strength of this process and the envelop mass.}

    As our primary observational constraint, we use a sample of subgiant stars from the \apogee{} \citep{apogee17} selected by the criteria in Roberts et al. (2023, in prep; see Appendix \ref{sec:jack}).
According to stellar evolution theory and observations \citep{gilroy89, korn+07, lind+08, souto+18, souto19}, these stars have not yet experienced first dredge up yet have well-mixed envelopes. So, subgiant surface abundances most accurately represent their birth composition. 
INTRO TO \apogee\ takes high-resolution near-infrared spectroscopic measurments of stars, allowing accurate determination of chemical abundances and stellar parameters. In particular, near-infrared C abundances are more reliable than in optical spectroscopy (CITE??).
Fig.~\ref{fig:subgiants} shows the subgiant sample plotted in \caah\ and \caafe{}. [C/Mg] increases with metallicity, and [C/Mg] decreases with [Mg/Fe] at fixed [Mg/H]. 
Using the subgiant abundance trends, we will develop a model of the enrichment sources and evolution of C.

@@

\section{Nucleosynthesis}

@@

We adapt the yield choices of elements besides C from \citet{james+21, james+23}.
Table \ref{tab:fiducial_mod} contains our fiducial yields. 
Following \citet{james+21, james+23}, we also take the \ia{} delay time distribution to be a $t^{-1.1}$ power-law with a minimum delay time of 140\,Myr, as suggested by the observations of \citet{maoz+12}.

@@

	\begin{tabular}{l l l l}
		\hline
        Element & $y^{\rm cc}$ & $\y^{\rm agb}$ & $y^{\rm ia}$ \\
		\hline
        C & Eq.~\ref{eq:zeta} & $2.9\times{\rm C11}$ &  0 \\
        O & 0.015 & 0 & 0 \\
        Mg & 0.00185 & 0 & 0 \\
        Fe & 0.0012 & 0 & 0.00214 \\
        N & 0.00072 & 0.0009$M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
		\hline
	\end{tabular}
\end{table}

@@

An \agb\ star is a low-mass ($\lesssim 8 M_{\sun}$) star during its final phase of evolution. Inside an \agb\ star, an inert CO core is surrounded by thermally-unstable He and H-burning shells below a convective envelope. 
In an \agb\ star, two competing processes determine the outcome of C production: \textit{third dredge up} and \textit{hot bottom burning}.  
Third dredge up accompanies thermal pulses in \agb\ stars, where material from the CO core is mixed with the envelope, increasing surface C abundances \citep{KL14}. The C yields of the star are increased as this C-enhanced envelope is released to the interstellar medium. 
Hot bottom burning is the activation of the CNO cycle\footnotemark{}
at the bottom of the convective envelope when $T\gtrsim 50\,{\rm MK}$. Because the $^{14}$N proton capture is the slowest component of the CNO cycle, the CNO cycle converts nearly all \C[12] into $^{14}$N \citep{solar-fusion}.

@@

Hot bottom burning and third dredge-up result in mass-dependent C yields. 
Stars less than \about{2}\,\Mo do not experience third dredge-up. As a result, these stars C abundances are only affected by first dredge-up, resulting in little change to C yields or slight destruction of C.
Between \about{2} and \about{5} \Mo, third dredge up becomes important, enriching the outer layers with C, making these stars the most important producers of C. 
In \agb\ stars more massive than \about{5}\,\Mo, both hot bottom burning and third dredge up occur; however, hot bottom burning is much more efficient, resulting in significant \C[12] destruction.

    In this work, we explore four different sets of \agb\ star yield tables from literature,
    providing necessary well-sampled grids in mass and metallicity. We refer to the yields from the following studies as the following,
\begin{description}
    \item \cxi: \citet{cristallo+11, cristallo+15}
    \item \kx: \citet{karakas10}
    \item \vxiii: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \kxvi: \citet{KL16, karakas+18}
\end{description}
In Appendix~\ref{sec:oob_models}, the masses, metallicites, and notable differences between each yield set are briefly described. 
For our models to match observations, we find that need to uniformly amplify these yield tables. We use \cxi\ table, amplified by a factor of 2.9, as the fiducial \agb\ yield.

@@

\begin{figure*}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}
        \caption[]{The net fractional \agb\ C yield  plotted as a function of initial stellar mass $M$ and colour-coded according to metallicity. The black dashed line shows $\y=0$ for reference. Each panel represents yields from one of four \agb\ models: \cxi, \kx{}, \vxiii{}, \kxvi{} (see Section \ref{sec:agb} and  Appendix \ref{sec:oob_models}). }
        \label{fig:y_agb}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics{y_agb_vs_t.pdf}
    \includegraphics{y_agb_vs_z.pdf}

    \caption[]{ C production by \agb\ stars as a function of \ssp{} age, normalized to the total mass $M_{\rm tot}$ produced at $t=10$\,Gyr. \textbf{Left:} The four \agb\ yield models from literature at solar metallicity (\cxi, \kx{}, \vxiii{}, or \kxvi{}). The dashed black line shows the delay time distribution of type Ia supernovae ($\propto t^{-1.1}$) for comparison. \textbf{Right:} The (\imf-weighted) \agb\ C yield $\Ycagb$ as a function of metallicity for each of the \agb\ yield models. ($\Ycagb$ is the net mass of C produced by \agb\ stars per unit mass of star formation, after 10\,Gyr and assuming a \citealt{kroupa01} \imf.) }

@@

Massive stars form $^{12}$C in their cores through the triple--$\alpha$ process.\footnote{The 3-$\alpha$ process fuses 3 He nuclei into O.} 
However, only C ejected through supernovae and stellar winds contributes to the yield. 
While there are many stellar models providing predictions of \cc{} yields, the results of these models are highly uncertain due to the complexity of stellar modeling.

@@

\begin{figure*}
    \centering
    \includegraphics{cc_yields.pdf}
    \caption[High-Mass Star Carbon Yields]{
        C yields from high-mass stars.
        \textbf{Left} The \imf-weighted \cc\ yield of C as a function of metallicity.
        \textbf{Right} The \cc\ [C/Mg] abundance ratio, defined in Eq.~\ref{eq:c_mg_cc}. The black line is the derived C yield from Section \ref{sec:equilibrium},
    $\Ycc = 0.0028 + 0.001 (Z/Z_{\odot})$. Yields are shown for tables from 
    \citet[red triangles]{WW95}, \citet[orange squares and diamonds]{sukhbold+16}, 
    \citet[green stars]{NKT13}, and \citet[blue circles]{LC18}. \citet{sukhbold+16} report yields for different black hole landscapes, while \citet{LC18} provide yields at different rotational velocities.
    In the top panel, the pink line denotes $\Ycagb$ from \cxi{} for comparison. All models include wind yields. 
    (OUT OF DATE)
}
    \label{fig:y_cc}
\end{figure*}

\section{Analytic Approximations}
In this section, we describe approximations of galactic abundance trends and C yield models. While these are insufficient to describe the galaxy in full detail, we can use these simplified models to inform our investigation.

@@

Galaxies, when moderated by metal-poor gas accretion and feedback-driven outflows, reach a chemical equilibrium. The production of new metals is balanced by losses to new stars and outflows \citep{larson72, dalcanton07, FD08, PS11, lilly13}.
While our galaxy is likely not in perfect equilibrium or described by a single, homogeneous chemical
envirnoment, the equilibrium approximation is nevertheless useful in understanding yields and
metallicity dependence of solar neighborhood stars \citep[e.g.][]{james_dwarf,james+23,WAF17}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,
%referee,
usenatbib]{mnras}

@@

% citations
\defcitealias{james+21}{J21}
\newcommand{\JJ}{\citetalias{james+21}}
\newcommand{\VICE}{\textsc{vice}}
\newcommand{\citetjack}{Roberts et al.~(2023, in prep.)}
\newcommand{\citepjack}{(Roberts et al.~2023, in prep.)}
\newcommand{\citealtjack}{Robert et al.~2023, in prep.}
\newcommand{\cxi}{\texttt{\hyperlink{C11}{C11}}}
\newcommand{\kx}{\texttt{\hyperlink{K10}{K10}}}
\newcommand{\kxvi}{\texttt{\hyperlink{K16}{K16}}}
\newcommand{\vxiii}{\texttt{\hyperlink{V13}{V13}}}
\newcommand{\cfactor}{2.3}

@@

% other
\newcommand{\ycmg}{\ensuremath{2.7 + 32\left(Z-\Zo\right)}}
\newcommand{\ycccmg}{1}
\newcommand{\yco}{1}
\newcommand{\fmeas}{20\%}

\makeatletter
\newcommand{\C}[1][\@nil]{
    \def\tmp{#1}%
    \ifx\tmp\@nnil%
        \ensuremath{\rm C}%
    \else%
        \ifmmode ^{#1}{\rm C}%
        \else $^{#1}$C%
        \fi%
\fi }
\makeatother

\newcommand{\Yct}{\ensuremath{y_{\rm C}}}
\newcommand{\Ycc}{\ensuremath{y_{\rm C}^{\rm cc}}}
\newcommand{\Yoc}{\ensuremath{y_{\rm Mg}^{\rm cc}}}
\newcommand{\Ycagb}{\ensuremath{y_{\rm C}^{\rm agb}}}
% \newcommand{\y}{\ensuremath{\rotatebox[origin=B,y=0.5ex]{180}{y}}}
\newcommand{\y}{Y}

\newcommand{\Mo}{%
    \ifmmode {\rm M}_{\sun}%
    \else {M$_{\sun}$}%
    \fi}
\newcommand{\Zo}{%
    \ifmmode Z_{\sun}%
    \else $Z_{\sun}$%
    \fi}

@@

% Abstract of the paper
\begin{abstract}
% context
C is an important element across astronomy; however, its origin remains poorly understood. 
% 
We aim to constrain the stellar yields of C through multi-zone Galactic chemical evolution models by comparing predictions with \apogee\ subgiants abundances.
% 
We find that \caafe\ is an emirical estimate of the delayed C sources, enabling us to estimate that \agb\ stars and \cc\ produce about 20\% and 80\% of C, respectively.  
The \caah\ trend instead represents the equilibrium abundances of C and Mg. 
We use the \caah\ trend to estimate the CCSNe C/Mg yield, determining that  $\Ycc/\Yoc = EQUATION$ when including \agb\ C. 
% misc
Our models are relatively independent of uniform scaling of yields and outflows, and alternate star formation histories. 
However, the stars which contribute to \agb\ C production and the \ia\ delay time distribution of Fe contribute uncertainties to our conclusions. 
% gas phase
While reliable gas-phase and low-metallicity measurments of C are challenging, we find that our model and a singlezone model with our recommended yields replicate the broad trends of \caah{} across different environments and metallicities.

@@

Carbon is a distinctive and well-studied element in astronomy. 
Formed in the cores of stars during He fusion, C is the lightest directly synthesized element after He, one of the only light elements formed in low-mass stars, and one the most abundant metals \citep[e.g.][]{jennifer19, KL14}.\footnotemark{}
C structurally changes the environments it pollutes -- regulating stellar evolution, facilitating the formation of stars and planets, and forming the basis of earthly life.
% Because the effects of first dredge up are mass-dependent, C and N abundances are frequently used as age indicators for RGB stars \citep{MG15, martig16, hasselquist19, vincenzo+21}. 
As such, understanding the origin of C has wide ranging implications. 
While we know both lower-intermediate-mass and high-mass stars produce C, the relative importance of each process is still unknown.

@@

We know, from observations of MW stars and (extragalactic) gas, that C/O traces a banana shape in O/H (see Fig.~\ref{fig:gas_phase} and section~\ref{sec:gas}). 
At the very lowest metallicities, where $[{\rm O/H}]\lesssim-2$,%
\footnote{In this paper, we use the standard notation for chemical abundances. $[A/B] = \log_{10}\left(A/B\right) - \log_{10}\left(A_{\sun}/B_{\sun}\right)$, i.e. $[A/B]$ is the logarithm of the ratio between A and B, scaled such that $[A/B]=0$ for the sun. Solar abundances are as measured in \citet{asplund+09}.}
%
observations suggest that C/O declines with metallicity.
From $-2 \lesssim {\rm [O/H]}\lesssim -1$, C/O is roughtly constant with metallicities. 
And, at higher metallicities $-1 \lesssim {\rm [O/H]}$, C/O increases with increasing metallicity.
    As our primary observational constraint, we use a sample of subgiant stars from the \apogee{} \citep{apogee17} selected by the criteria in Roberts et al. (2023, in prep).
    According to stellar evolution theory and observations \citep{gilroy89, korn+07, lind+08, souto+18, souto19}, these stars have not yet experienced \textit{first dredge up}%
\footnote{First dredge up is when material from CNO processed core is mixed with the (convective) stellar envelop during the ascent onto the RGB. As a result, C is decreased and N is increased depending on the strength of this process and the envelop mass.}
yet have well-mixed envelopes. So, subgiant surface abundances most accurately represent their birth composition. 
In contrast, evolved RGB stars require modelling-dependent corrections to recover birth C abundances \citep[e.g.][]{vincenzo+21}. 
Fig.~\ref{fig:subgiants} shows the subgiant sample plotted in \caah\ and \caafe{}. [C/Mg] increases with metallicity, and [C/Mg] decreases with [Mg/Fe] at fixed [Mg/H]. 
Using the subgiant abundance trends, we will develop a model of the enrichment sources and evolution of C.

@@

Galactic chemical evolution (\gce) is a powerful tool, capable of uncovering the origins of the elements. 
Each enrichment process has characteristic chemical signatures and timescales, enabling us to reconstruct chemical histories.
Many previous works have used GCE in attemt to understand C abundances, whether through using theoretical stellar models \citep{DTS78, prantzos+18, chiappini+03} 
or understanding  observation (\citealt{tinsley79, HEK00, BF06, rybizki+17, berg+19, KKL20};
See also review in \citealt{romano22}).
%
Every study agrees that C is produced by a combination of high-mass and \lims{}, different studies disagree on which process is dominant. 
For example CITES conclude \hms{} dominate wherease CITES conclude \lims{} contribute the majority of C.
C is also generally understood to have strong metallicity dependent CCSNe enrichment.

One of the primary uncertainties of GCE models are nucleosynthetic yields. Yield predictions -- the amounts of each chemical element stars produce --
are shaped by poorly understood processes, including mass loss, nuclear reaction rates, rotational mixing, convection, and explodability \citep{romano+10,KL14,ventura+13, LC18, emily+21}.
To better understand where C comes from and how it evolves, our aim is to combine \apogee\ observations and multi-zone models to develop observationally-consistant yields.
\cite{james+23} examined similar \gce{} models of N (which is closely related to C), finding that trends in N and O are explained by the metallicity dependence of N/O yields. \citet{james+23} determine that \agb\ N abundances roughly depend linearly on metallicity (i.e. $y_{\rm N}/y_{\rm O} \propto Z$). 
Here, we extend their models to C, deriving similar constraints on C/Mg yields. We assess which yield prescriptions reproduce Galactic abundance trends while investigating the impact of \gce{} model assumptions, such as the star formation history (\sfh{}) and outflow mass loading.

\section{Nucleosynthesis}
We adapt the yield choices of elements besides C from \citet{james+21, james+23}.
Table \ref{tab:fiducial_mod} contains our fiducial yields. 
Following \citet{james+21, james+23}, we also take the \ia{} delay time distribution to be a $t^{-1.1}$ power-law with a minimum delay time of 140\,Myr, as suggested by the observations of \citet{maoz+12}.

@@

@@

	\begin{tabular}{l l l l}
		\hline
        Element & $y^{\rm cc}$ & $\y^{\rm agb}$ & $y^{\rm ia}$ \\
		\hline
        C & Eq.~\ref{eq:zeta} & $2.3\times$\cxi &  0 \\
        O & 0.015 & 0 & 0 \\
        Mg & 0.00185 & 0 & 0 \\
        Fe & 0.0012 & 0 & 0.00214 \\
        N & 0.00072 & 0.0009$M\left(\frac{Z}{Z_\odot}\right)$ & 0\\
		\hline
	\end{tabular}
\end{table}

@@

An \agb\ star is a low-mass ($\lesssim 8\,\Mo$) star during its final phase of evolution. Inside an \agb\ star, an inert CO core is surrounded by thermally-unstable He and H-burning shells below a convective envelope. 
In an \agb\ star, two competing processes determine the outcome of C production: \textit{third dredge up} and \textit{hot bottom burning}.  
Third dredge up accompanies thermal pulses in \agb\ stars, where material from the CO core is mixed with the envelope, increasing surface C abundances \citep{KL14}. The C yields of the star are increased as this C-enhanced envelope is released to the interstellar medium. 
Hot bottom burning is the activation of the CNO cycle\footnotemark{}
at the bottom of the convective envelope when $T\gtrsim 50\,{\rm MK}$. Because the $^{14}$N proton capture is the slowest component of the CNO cycle, the CNO cycle converts nearly all \C[12] into $^{14}$N \citep{solar-fusion}.

@@

Hot bottom burning and third dredge-up result in mass-dependent C yields. 
Stars less than \about{2}\,\Mo do not experience third dredge-up. As a result, these stars C abundances are only affected by first dredge-up, resulting in little change to C yields or slight destruction of C.
Above \about{2}\,\Mo{}, third dredge up becomes important, enriching the outer layers with C.
\agb\ stars between 2 and 5 \Mo are the most abundant producers of C.
In \agb\ stars more massive than \about{5}\,\Mo, both hot bottom burning and third dredge up occur; however, hot bottom burning is much more efficient, resulting in significant \C[12] destruction.

    In this work, we explore four different sets of \agb\ star yield tables from literature,
    providing necessary well-sampled grids in mass and metallicity. We refer to the yields from the following studies as the following,
For our models to match observations, we find that need to uniformly amplify these yield tables. We use \cxi\ table, amplified by a factor of 2.9, as the fiducial \agb\ yield.
Variations in models are due to different treatments of reaction rates, convection, and mass-loss.
\begin{description}
    \item \hypertarget{C11}{\texttt{C11}}: \citet{cristallo+11, cristallo+15}
    \item \hypertarget{K10}{\texttt{K10}}: \citet{karakas10}
\item \hypertarget{V13}{\texttt{V13}}: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \hypertarget{K16}{\texttt{K16}}: \citet{KL16, karakas+18}
\end{description}

@@

\begin{table*}
	\centering
    \caption[]{For each \agb\ yield set, the \imf-averaged \agb\ C yield at solar metallicity $y_{\rm C, 0}^{\rm agb}$ and the multiplicative factor reaches an \agb\ contribution of 20\% $\alpha_{\rm agb, 20}$.}

	\label{tab:alpha_agb}
    \begin{tabular}{c ccc p{5cm} p{5cm}} % four columns, alignment for each
		\hline 
        \agb\ table 
                & $y_{\rm C}^{\rm agb}(\Zo)$ & $\zeta^{\rm agb}(\Zo)$ & $\alpha_{20}^{\rm agb}$
                & masses (\Mo) & metallicites 
                \\
        \hline
        \cxi 
                &  0.00042 & -0.0175 & 2.4
                & 1.3, 1.5, 2, 2.5, 3, 4, 5, 6
                & 0.0001, 0.0003, 0.001, 0.002, 0.003, 0.006, 0.008, 0.01, 0.014, 0.02
                \\
        \kx 
                & 0.00064 & -0.059 & 1.6
                & 1, 1.25, 1.5, 1.75, 1.9, 2.25, 2.5, 3, 3.5, 4, 4.5, 5, 5.5, 6 
                &  0.0001, 0.004, 0.008, 0.02
                \\
        \vxiii 
                & 0.00022 & -0.021 & 4.5
                & 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 6, 6.5, 7
                & 0.0003, 0.001, 0.002, 0.004, 0.008, 0.014, 0.04
                \\
        \kxvi 
                & 0.0005 & -0.029 & 2.0
                & 1, 1.25, 1.5, 1.75, 2.25, 2.5, 2.75, 3, 3.25, 3.5, 3.75, 4, 4.5, 5, 5.5, 6, 7 
                & 0.0003, 0.001, 0.002, 0.004, 0.008, 0.014, 0.04
                \\
		\hline
	\end{tabular}
\end{table*}

\begin{figure*}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}
        \caption[]{The net fractional \agb\ C yield  plotted as a function of initial stellar mass $M$ and colour-coded according to metallicity. The black dashed line shows $\y=0$ for reference. Each panel represents yields from one of four \agb\ models: \cxi{}, \kx{}, \vxiii{}, \kxvi{}, and our analytic model (see sections \ref{sec:agb}) }

        \label{fig:y_agb}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics{y_agb_vs_z.pdf}
    \includegraphics{y_agb_vs_t.pdf}

    \caption[]{ C yields from \agb\ stars as a function of \ssp{} age, as a fraction of the C yield at $t_{\rm end}=10\,$Gyr. 
\textbf{Left} The (\imf-weighted) \agb\ C yield $\Ycagb$ as a function of metallicity for each of the \agb\ yield models. ($\Ycagb$ is the net mass of C produced by \agb\ stars per unit mass of star formation, after 10\,Gyr and assuming a \citealt{kroupa01} \imf.)
    \textbf{Right} Our four considered \agb\ yield models at solar metallicity (\cxi{}, \kx{}, \vxiii{}, \kxvi{}. The dashed red line shows the delay time distribution of type Ia supernovae ($\propto t^{-1.1}$) for comparison, and the minimum of V13 is $y(t)/y(t_{\rm end}=-3.3$. 
}

@@

Massive stars form $^{12}$C in their cores through the triple--$\alpha$ process.\footnote{The 3-$\alpha$ process fuses 3 He nuclei into \C[12].} 
However, only C ejected through supernovae and stellar winds contributes to the yield. 
While there are many stellar models providing predictions of \cc{} yields, the results of these models are highly uncertain due to the complexity of stellar modeling.

@@

\begin{figure*}
    \centering
    \includegraphics{cc_yields.pdf}
    \caption[]{
        C yields from high-mass stars.
        \textbf{Left} The \imf-weighted \cc\ yield of C as a function of metallicity.
        \textbf{Right} The \cc\ [C/Mg] abundance ratio, defined in Eq.~\ref{eq:c_mg_cc}. The black line is the derived C yield from Section \ref{sec:equilibrium} and Eq.~\ref{eq:zeta}. Yields are shown for tables from 
    \citet[red triangles]{WW95}, \citet[orange squares and diamonds]{sukhbold+16}, 
    \citet[green stars]{NKT13}, and \citet[blue circles]{LC18}. \citet{sukhbold+16} report yields for different black hole landscapes, while \citet{LC18} provide yields at different rotational velocities.
    In the top panel, the pink line denotes $\Ycagb$ from \cxi{} for comparison. All models include wind yields. 
}
    \label{fig:y_cc}
\end{figure*}

@@

Here, we assume a simple \textit{one-zone} chemical evolution model \cite[e.g.][]{tinsley80, pagel09, matteucci21}.
Galaxies, when moderated by metal-poor gas accretion and feedback-driven outflows, reach a chemical equilibrium. The production of new metals is balanced by losses to new stars and outflows \citep{larson72, dalcanton07, FD08, PS11, lilly13}.
While our galaxy is likely not in perfect equilibrium or described by a single, homogeneous chemical
envirnoment, the equilibrium approximation is nevertheless useful in understanding yields and
metallicity dependence of solar neighborhood stars \citep[e.g.][]{james_dwarf,james+23,WAF17}.

\cite{james+23} show that trends in N and O abundance ratios are set by the yield ratio and their metallicity dependences. 
Here, we find a similar conclusion for C and Mg, where equilibrium C/Mg is given by
\begin{equation}\label{eq:equilibrium_yields}
    \frac{Z_{\rm C}^{\rm eq}}{Z_{\rm Mg}^{\rm eq}} = \frac{\Ycc + \Ycagb }{y_{\rm Mg}}.
\end{equation}

In Fig.~\ref{fig:analytic}, we show the inferred total C yields, based on this equation, and our best fitting linear model. From a linear regression, we suggest that
\begin{subequations}\label{eq:yc_inferred}
    \begin{align}
        \frac{\Yct(Z)}{y_{\rm O}} &\approx \frac{1}{3} + 4 \left(Z-\Zo\right) \\
        \frac{\Yct(Z)}{y_{\rm Mg}} &\approx 2.7 + 32 \left(Z-\Zo\right).
    \end{align}
\end{subequations}
These yield ratios results in an equilibrium abundance $[{\rm C}/\alpha] = -0.09$, which is consistent with the subgiant sample and is within \about{20\%} of the solar C/Mg mixture from \citet{asplund+09}.

Both observational and theoretical uncertainties limit the accuracy of our relative yield predictions.
As the metallicity range of our data is small ($-0.4\lesssim {\rm [Mg/H]} \lesssim 0.4$),
our model does not extend beyond near-solar metallicities.
Additionally, the derived yields will be systematically biased if the galaxy is out of equilibrium, for example, due to a recent starburst \citep{mor+19,isern19}. 
As we will discuss in Section \ref{sec:outflows}, the normalization of yields and $\eta$ is degenerate. If we increase yields and $\eta$ by a similar factor, equilibrium abundances from Eq.~\ref{eq:equilibrium_yields} would be unchanged, so our model cannot distinguish changes in both yields and outflows.

With the total C yield of Eq.~\ref{eq:yc_inferred} and given an \agb\ C yield, we can derive an observationally-consistant \cc\ C yield.
\begin{subequations}
    \begin{gather}
        \Ycc(\Zo) = \Yct(\Zo) - \Ycagb(\Zo)\\
        \zeta^{\rm cc} \equiv \frac{d\Ycc(\Zo)}{dZ} = \frac{d\Yct(\Zo)}{dZ} - \frac{d\Ycagb(\Zo)}{dZ}
    \end{gather}
\end{subequations}
This method reduces the number of free parameters of the model and enables all models to match \caah\ trends in \apogee{} subgiants when considering \agb\ yields with different metallicity dependences and normalizations. 
Table~\ref{tab:alpha_agb} contains our adaptive (regression-derived) values of $\zeta^{\rm agb}$ and $\Ycagb(\Zo)$. 
For example, in our fiducial model, $\Ycc = 0.004 + 0.102 (Z-\Zo)$.
Our full \cc\ C yield model adds a low-metallicity enhancement term, which is insignificant at solar metallicities (e.g. thin disk stars) but enables models in Section~\ref{sec:gas} to match low-metallicity environments.
We parameterize low-metallicity enrichment with $y_l$ and $Z_l$, the yield and transition metallicity for enhanced low-metallicity yields.
All-together, 
\begin{equation}\label{eq:zeta}
    \Ycc = y_{\rm C, 0}^{\rm cc} + \zeta^{\rm cc}\,(Z-\Zo) + \frac{2\,y_l}{1 + Z/Z_l}.
\end{equation}

\begin{figure}
    \centering
    \includegraphics{analytic.pdf}
    \caption[]{Inferred total C yields as a function of metallicity. We assume chemical equilibrium (orange curve, see discussion in Section \ref{sec:equilibrium}). Blue points are the median value of $\Ycc$ for each (number) bin in [Mg/H] with uncertainties based on the 16--84 percentile range.
    }
    \label{fig:analytic}
\end{figure}
